Practical Machine Learning Course Project
=============================

##Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

##Process the data

```{r}
library(caret)
library(dplyr)
library(randomForest)
library(e1071)
pmltrain <- read.csv("pml-training.csv",na.strings=c("NA","","#DIV/0!"))
##split pmltrain into train set and test set
inTrain <- createDataPartition(y = pmltrain$classe, list = FALSE, p=0.7)
trainData <- pmltrain[inTrain,]
testData <- pmltrain[-inTrain,]
```

##Select predictors

```{r}
##identify variables that are mostly NAs
naprops <- colSums(is.na(trainData))/nrow(trainData)
mostlyNAs <- names(naprops[naprops > 0.75])
mostlyNACols <- which(naprops > 0.75)
##take a random small sample from the train set
set.seed(1256)
smalltrain <- trainData %>% tbl_df %>% sample_n(size=1000)
##remove the variables that are mostly NAs
smalltrain <- smalltrain[,-mostlyNACols]
##remove row number and user name as candidate predictors
smalltrain <- smalltrain[,-grep("X|user_name",names(smalltrain))]
##remove the cvtd_timestamp variable as a candidate predictor
smalltrain <- smalltrain[,-grep("cvtd_timestamp",names(smalltrain))]
##remove candidate predictors that have near zero variance
smalltrain <- smalltrain[,-nearZeroVar(smalltrain)]
##list the candidate predictors
modelVars <- names(smalltrain)
modelVars1 <- modelVars[-grep("classe",modelVars)]
modelVars1
```

##Build a random forest model

```{r}
set.seed(57)
cleanedTrainData <- trainData[,modelVars]
modelFit <- randomForest(classe ~., data=cleanedTrainData, type="class")
##estimate the errors
predTrain <- predict(modelFit,newdata=trainData)
###in sample error
confusionMatrix(predTrain,trainData$classe)$table
###our of sample error
classe_col <- grep("classe",names(testData))
predTest <- predict(modelFit, newdata = testData[,-classe_col], type="class")
confusionMatrix(predTest,testData$classe)
```

This model has an out of sample accuracy of 0.998.

##Prediction using the model

```{r}
pmltest <- read.csv("pml-testing.csv",na.strings=c("NA","","#DIV/0!"))
predplmtest <- predict(modelFit, newdata = pmltest, type="class")
predplmtest
```

